### 2022-04-21 initialdraft.Rmd
Cat's graphs pasted in

---
title: 'finalproj'
author: "Jocelyn, Ana, Cat"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(lubridate) 
library(ggthemes)
library(plotly)
library(ggdark)
```

```{r}
healthy_data <- read_csv("healthy_lifestyle_city_2021.csv", 
    col_types = cols(Rank = col_number(), 
        `Sunshine hours(City)` = col_number(), 
        `Cost of a bottle of water(City)` = col_number(), 
        `Obesity levels(Country)` = col_number(), 
        `Life expectancy(years) (Country)` = col_number(), 
        `Pollution(Index score) (City)` = col_number(), 
        `Annual avg. hours worked` = col_number(), 
        `Happiness levels(Country)` = col_number(), 
        `Outdoor activities(City)` = col_number(),
        `Number of take out places(City)` = col_number(), 
        `Cost of a monthly gym membership(City)` = col_number()))

# converting to dollars
healthy_data_dollars <- healthy_data %>% 
  mutate(water_bottle_cost = `Cost of a bottle of water(City)`*1.3,
         gym_cost = `Cost of a monthly gym membership(City)`*1.3)
```

> Because economic performance is a common standard for measuring the success of a given country, we chose to investigate if there is a correlation between gross domestic product (GDP), which is a measure of the total volume of goods and services produced within in country in a given year, and happiness levels as calculated by the **survey**. We used population data to calculate GDP per capita (thereafter, GDPPC) so as to control for differences in population between countries. We believe that GDPPC provides a helpful, though imperfect measure of the standard of living within a country. GDPPC operates with the assumption that every individual within a country gets an equal slice of the GDP pie, which is unrealistic, however, this is an assumption we will make for the purpose of this analysis. We see that countries with higher GDPPC tend to correlate with higher happiness levels than countries with lower GDPPC. We wondered if countries with a higher GDPPC would be more unhappy reflecting that they may work longer hours in order to have such a high level of production. It is possible that rather than hours worked, GDP reflects differences in technology, production processes, management, or human capital.

[GDP data source](https://worldpopulationreview.com/countries/countries-by-gdp)

```{r}
country_GDP <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQL_6se7aBLQSP92RgP5DkvQ_quTyzsTxskf9ZFpvE_5ifMS5whQ1Lr0JRNMtd1ZGAaVXlEEMECPMHr/pub?output=csv", 
    col_types = cols(GDPM = col_number(), 
        GDP = col_number(), pop = col_number(), 
        GDPPC = col_number()))

healthy_data_2 <- healthy_data_dollars %>% 
  left_join(country_GDP,
            by = c("City" = "city"))

options(scipen=999)

ggplot(healthy_data_2, aes(y = fct_reorder(country, GDPPC), x = GDPPC, color = `Happiness levels(Country)`)) +
  geom_point() +
  labs(title = "Gross Domestic Product per Capita by Country and Happiness Level",
       caption = "Data from: ADD LATER",
       x = "GDP per capita (USD)",
       y = "",
       color = "Happiness Level 
       (1-10 scale)") +
  theme(plot.title = element_text(hjust = 1)) +
  theme_classic() 
```

#Starting to mess with hours worked but realized that missing data is going to be a problem
```{r}
ggplot(healthy_data_2, aes(y = fct_reorder(City, `Hours Worked`), x = `Hours Worked`, color = `Happiness levels(Country)`)) +
  geom_point() +
  labs(title = "",
       caption = "Data from: ADD LATER",
       x = "",
       y = "",
       color = "Happiness Level 
       (1-10 scale)") +
  theme(plot.title = element_text(hjust = 1)) +
  theme_classic() 
```

Cat's graph(s):

We chose three variables that we think are important contributing factors to overall happiness, influenced of course by where we are and salient factors we feel attuned to. Our variables, life expectancy, sunshine hours, and hours worked, are compared against overall happiness for each place to see if there is a correlation between more or less of each variable. We are also keeping in mind intangible factors such as state of mind and the impacts of cultural attitudes.

```{r, r.height = 20, r.width = 15, fig.alt= "This dot plot compares the level of happiness to the sunshine hours in 43 different locations from the Healthy Lifestyle Cities Report 2021. The dots are all assigned a color based on location, and plotted at each discrete point. There appears to be no strong correlation between isolating the sunshine hours and comparing them to the happiness level. More factors combine to make up overall happiness."}
healthy_lifestyle_data_cat <- healthy_data_dollars[-c(20),]

healthy_lifestyle_data_cat2 <- healthy_lifestyle_data_cat %>% 
  select(`City`, `Happiness levels(Country)`, `Sunshine hours(City)`) %>%

  ggplot(aes(x = `Happiness levels(Country)`, 
             y = factor(`Sunshine hours(City)`), 
             color = (`City`))) +
  geom_count() +
  scale_y_discrete(breaks = c(1405, 1884, 2363, 2983, 3542))+

  labs(title = "Is there a correlation between sunshine hours and happiness levels?",
       x = "Happiness Level",
       y = "Sunshine Hours",
       caption = "Graph By Cat Terres; data sourced from Kaggle, 
       'Healthy Lifestyle Cities Report 2021'") +

  theme_foundation() +
  theme(legend.title=element_blank())+
  theme(legend.background = element_blank())

```

plotly for cities
```{r}
ggplotly(healthy_lifestyle_data_cat2,
         tooltip = c("text", 
                     "x",
                     "(City)"))
```

Life expectancy vs. Happiness (factor in city)
Does happiness affect life expectancy?

```{r, fig.alt= "For this graph, the cities have been placed into bins based on the happiness level and the range they fall into. For example, a country with a happiness level of 4.5 would be placed within the box plot above the "4-5" on the x axis. They are then plotted using upper and lower quartiles to show the distribution of life expectancies."}
bin_size <- 1
healthy_data_dollars %>% 
  select("City" = `City`, "Happiness Level" = `Happiness levels(Country)`, "Life Expectancy" = `Life expectancy(years) (Country)`) %>% 
  arrange(desc(`Life Expectancy`)) %>%

  mutate(bin_dist = paste(as.character(as.integer(`Happiness Level`)),as.character(as.integer(`Happiness Level`+1)), sep = "-")) %>%  
  ggplot() +
  geom_boxplot(aes(x = factor(bin_dist), 
         y = `Life Expectancy`,
         outlier.color = NULL,
         fill = bin_dist
         )) +
  labs(title = "Does reported greater happiness have a positive effect on life expectancy?",
       x = "Happiness Level",
       y = "Life Expectancy",
       fill = "") +

  guides(y = guide_axis(angle = 45))

```

Hours Worked vs. Happiness (factor in city)
Is there a correlation between happiness and hours worked?
#This one is still a work in progress
```{r}
##waiting on revised hours worked data
healthy_data_dollars %>% 
  select(`City`, "Happiness Level" = `Happiness levels(Country)`, "Average Annual Hours Worked" = `Annual avg. hours worked`) %>% 
  arrange(desc(`Average Annual Hours Worked`)) %>%

  mutate(bin_dist = paste(as.character(as.integer(`Happiness Level`)),as.character(as.integer(`Happiness Level`+1)), sep = "-")) %>%

  ggplot(aes(x = factor(bin_dist),
             y = `Average Annual Hours Worked`
             )) +
  geom_col()

```

Joce's graph Which cities are the sunniest and least sunny?
```{r, fig.alt = "Graph of the top 10 and bottom 10 cities by annual sunshine hours.The cities are mapped on the y-axis while the sunshine hours is mapped on the x-axis. Cairo, Tel Aviv, and Los Angeles were the top three trending just over 3000 annual sunshine hours while Brussels, Dublin, and Taipei were the bottom three with around 1500 annual sunshine hours."}
top_10 <- healthy_data %>% 
  arrange(desc(`Sunshine hours(City)`)) %>% 
  slice(1:10) %>% 
  ggplot(aes(y = fct_reorder(City,`Sunshine hours(City)`), x = `Sunshine hours(City)`)) +
  geom_col(fill = "khaki") +
  labs(title = "Top 10 and Bottom 10 Cities by Annual Sunshine Hours",
       y = "",
       x = "") +
  dark_theme_minimal()

bottom_10 <- healthy_data %>% 
  arrange(`Sunshine hours(City)`) %>% 
  slice(1:10) %>% 
  ggplot(aes(y = fct_reorder(City,`Sunshine hours(City)`), x = `Sunshine hours(City)`)) +
  geom_col(fill = "khaki") +
  xlim(0,3540) +
  labs(y = "",
       x = "",
       caption = "Data from UK Healthy Lifestyle Cities Report 2021, Graph by Jocelyn Radke") +
  dark_theme_minimal()

grid.arrange(top_10, bottom_10, nrow=2)
```

Ana's graph

```{r, fig.alt= "Scatterplot showing the relationship between the precentage of Obesity in each city's respective country and the number of take out places in that city." }
healthy_data %>% 
  head(n =10) %>% 
  select(City, `Obesity levels(Country)`, `Number of take out places(City)`) %>% 
  ggplot(aes(x = `Obesity levels(Country)`, y = `Number of take out places(City)`, color = City))+
  geom_point(aes(color = City), size = 3)+
  labs(title= "Obesity Percentage Variation with Increasing Number of Take Out Places", x = "Country-wide Obesity Levels", y = "Number of Take Out Places in Each City")+
  scale_color_brewer(palette="Paired"
                     )+
  theme(
  panel.background = element_rect(fill = "steelblue4",
                                colour = "steelblue4",
                                size = 0.5,
                                ),
  panel.border = element_blank())

```